Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-15 doc/report/msc-report.tex
doc/report/msc-report.tex: WIP before pub.

\section{Background}

@@

\section{Problem description}
% Problemsbeskrivning.

@@

\section{Goal}
% Syftesformulering.

@@

\section{Scope}
% Avgransning.

@@

\section{Related work}

@@

\section{Overview}

@@

% CONCEPTS IN DATA WAREHOUSING, by clark de jean
\chapter{Concepts in Data Warehousing}

\chapterquote{Some random quote might be suitable here.}{Per}

@@

This chapter introduces data warehouses and the concept of data cubes. It then
introduces common operations on data cubes: projections, slicing and dicing,
and aggregation.

@@

At the core of any OLAP system is the OLAP cube (also known as a
multi-dimensional cube or a hypercube). The cube consists of numeric facts
-- called measures -- which are grouped by dimensions. For example, the number
of nails sold is a measure while country, month and type of nail are
dimensions. It is from OLAP that [we] have appropriated the concept of a data
cube.

\section{Data cubes}

@@

\subsection*{Projections}

@@

For example, consider the three-dimensional data cube that represents sales of
nails; the dimensions are nail type, country and month. A projection from that
cube to one with two dimensions -- month and nail type -- would flatten the
country dimension: the measurements in the new cube represent the global sales
of each nail type for each month.

\subsection*{Slice and dice}

@@

\section{Aggregates}

@@

@@

This chapter introduced Data Warehouses in general and OLAP in particular. The
concept of data cubes was introduced by OLAP. The chapter was rounded off with
a discussion of dimensionality reducing operations and aggregations of data
cubes.

@@

Introduce the chapter here.

@@

@@

We explored the problem domain by asking questions about the environment, the
system we were to build and its requirements.

@@

% this is a little thin. The prototypes are small pieces of code that would
answer questions such as Is it at all possible to.. ? and What is the most
efficient way to..?

@@

\subsection*{Literature studies}
% vet inte riktigt vad jag vill saga om det har.

\subsection*{Summary}

This section has talked about the procedure we followed during our thesis work. It described our exploratory model and the prototyping we did, as well as the design phase that led up to the implementation.

@@

Hadoop is a free software implementation of Google's MapReduce-framework; it is
written in Java and is part of the Apache Software Foundation. Hadoop is a
fairly mature piece of software and is used in production at numerous companies,
among them Last.fm.
\cite{hadoop}

@@

HDFS is a free software implementation of the GFS \cite{gfs}.

@@

\subsection{Data model}

\begin{table}[h]
    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
        \hline
        \small \bf Row Key & \small \bf Time & \small \bf Column &
        \multicolumn{2}{|c|}{\small \bf Column} \\
         & \small \bf Stamp & \small "contents:" &
        \multicolumn{2}{|c|}{\small "anchor:"} \\
        \hline
         & t9 & & \small "anchor:cnnsi.com" & "CNN" \\
         & t8 & & \small "anchor:my.lok.ca" & "CNN.com" \\
        "com.cnn.www" & t6 & \small "{\textless}html\textgreater..." & & \\
         & t5 & \small "{\textless}html\textgreater..." & & \\
         & t3 & \small "{\textless}html\textgreater..." & & \\
        \hline
        \end{tabular}
        \caption{Example HBase table.}
    \end{center}
\end{table}

@@

\section{Python}

Because there existed a Python module for writing MapReduce programs (Dumbo),
Python was a natural choice during the design and exploration phases. The main
reason was the rapid development of Python programs, several times shorter than
their Java counter parts made them faster to develop.

@@

Summarize the method chapter.

Working methods.

Tools.

Something else-or-another.

@@

\begin{lstlisting}[language=Python]
   def map(key, value):
      log = parse_apache_log(value)
      dimensions = {'useragent' : log.useragent,
                    'path'      : log.path,
                    'status'    : log.statuscode}
      values = {'pageviews' : 1,
                'bytes'     : log.size}

@@

- Something about cardinality -

@@

-- Limits on cardinality --

@@

-- The implementation of all that --

@@

The reducer needs to communicate to the OutputReader under what rowkey,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form
\texttt{\{"column-family:qualifier" : value\}}.

@@

The data export server is built with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URI it was fed. \cite{python_paste}

@@

Middleware which dispatches to applications based on the URI. The URLParser
removes the first part of the URI and tries to find a matching application
to relay the request to. For instance, consider the URI
\texttt{/client/file.html}. The first part---\texttt{client}---will be
removed and the request \texttt{/file.html} will be sent to the application
named \texttt{client.py}.

@@

Dispatches request to serve static files from the project's
\texttt{clients} directory. Custom clients---which interact with Zohmg---can
be built with HTML and JavaScript, see the examples.

@@

% ?
Fetches data from HBase via its Thrift interface. The data is served to the
user as JSONP\footnote{JSON with padding: JSON extension which adds name of
callback function as an argument to the function itself.}.

@@

@@

\section{Installation}

@@

\section{Deployment}

@@

\section{Import}

@@

\section{Export}

@@

\noindent The above query would return the pageviews for SE between 1 January
2009 and 4 April 2009.

@@

>>>>

\section*{Background}

@@

\section*{Problem description}
% Problemsbeskrivning.

@@

\section*{Goal}
% Syftesformulering.

@@

\section*{Scope}
% Avgransning.

@@

\section*{Related work}

@@

\section*{Overview}

@@

% CONCEPTS IN DATA WAREHOUSING, by clark de jean
\chapter{Concepts}

\chapterquote{ I don't let go of concepts - I meet them with understanding.
Then they let go of me.}{Byron Katie}

@@

This chapters first section presents data warehouses and the concept of data
cubes. It then introduces common operations on data cubes: projections, slicing
and dicing, and aggregation.

Secondly, this chapter features new techniques in distributed storage and
computing; these techniques are Google's BigTable and MapReduce.

@@

At the core of any OLAP system is the OLAP cube (also known as a
multi-dimensional cube or a hypercube). The cube consists of numeric
facts---called measures---which are grouped by dimensions. For example, the
number of nails sold is a measure while country, month and type of nail are
dimensions. It is from OLAP that [we] have appropriated the concept of a data
cube.

\subsection*{Data cubes}

@@

\subsubsection*{Projections}

@@

For example, consider the three-dimensional data cube that represents sales of
nails; the dimensions are nail type, country and month. A projection from that
cube to one with two dimensions---month and nail type---would flatten the
country dimension: the measurements in the new cube represent the global sales
of each nail type for each month.

\subsubsection*{Slice and dice}

% TODO: this is actually exactly the same thing as projection.

@@

\subsection*{Aggregates}

@@

\section{New techniques}

Classic data warehouses use relational databases as storage. A star or snowflake
schema would be used to modell the dimensions.

New techniques that have appeared include column-oriented, schema free
databases. Google's BigTable is such a database. Another new technique for
distributed processing is Google's MapReduce.

\subsubsection{Data model}

% TODO: fix positioning.
\begin{table}[h]
    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
        \hline
        \small \bf Row Key & \small \bf Time & \small \bf Column &
        \multicolumn{2}{|c|}{\small \bf Column} \\
         & \small \bf Stamp & \small "contents:" &
        \multicolumn{2}{|c|}{\small "anchor:"} \\
        \hline
         & t9 & & \small "anchor:cnnsi.com" & "CNN" \\
         & t8 & & \small "anchor:my.lok.ca" & "CNN.com" \\
        "com.cnn.www" & t6 & \small "{\textless}html\textgreater..." & & \\
         & t5 & \small "{\textless}html\textgreater..." & & \\
         & t3 & \small "{\textless}html\textgreater..." & & \\
        \hline
        \end{tabular}
        \caption{Example HBase table.}
    \end{center}
\end{table}

\begin{lstlisting}[caption=BigTable data model illustration.,captionpos=b]
   (row key,column key,timestamp) -> value
\end{lstlisting}

@@

This chapter introduced Data Warehouses in general and OLAP in particular. The
concept of data cubes was introduced by OLAP. Dimensionality reducing operations
and aggregations of data cubes was also introduced in this chapter.

The chapter rounded of with a presentation of new techniques for distributed
storage and processing; namely Google's BigTable and MapReduce techniques.

@@

The first section in this chapter will present the work procedure during the
master thesis. The following chapters will discuss selected tools and their
impact on design and implementation choices for the Zohmg system.

@@

% TODO: generally use passive tense.
% TODO: renskriv.

@@

The problem domain was explored by asking, the system to be built and the
requirements upon it.

@@

% this is a little thin. The prototypes are small pieces of code that would
TODO: answer questions such as Is it at all possible to.. ? and What is the most
efficient way to..?

\subsubsection*{Literature studies}
% vet inte riktigt vad jag vill saga om det har.

\subsubsection*{Python}

Because there existed a Python module for writing MapReduce programs (Dumbo,
further later in this chapter), Python was a natural choice during the design
and exploration phases. The main reason was the rapid development of Python
programs, several times shorter than their Java counter parts made them faster
to develop.

@@

@@

Hadoop is a free software implementation of Google's MapReduce-framework; it is
written in Java, and is part of the Apache Software Foundation. Hadoop is a
fairly mature piece of software and is used in production at numerous companies,
among them Last.fm. \cite{hadoop}

@@

HDFS is a free software implementation of the GFS\footnote{Google File System}.
\cite{gfs}

@@

@@

@@

The first section of this chapter talked about the procedure during the thesis
work. It described the exploratory model and the prototyping performed, as well
as the design phase that led up to the implementation.

The following sections presented and discussed the selected tools; they also
discussed the impact on design and implementation choices.

@@

\begin{lstlisting}[language=Python,caption=An example map function.,captionpos=b]
   def map(key, value):
      log = parse_apache_log(value)
      dimensions = {'useragent' : log.useragent,
                    'path'      : log.path,
                    'status'    : log.statuscode}
      values = {'pageviews' : 1,
                'bytes'     : log.size}

@@

% TODO: analysis.
- Something about cardinality -

@@

% TODO: method.
-- Limits on cardinality --

@@

% TODO: implementation.
-- The implementation of all that --

@@

% TODO: data interchange.
The reducer needs to communicate to the OutputReader under what rowkey,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form
\texttt{\{"column-family:qualifier" : value\}}.

@@

The data export server is built with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URL it was fed. \cite{python_paste}

@@

The URLParser class is middleware that comes with Python Paste; its purpose is
to dispatch to applications based on the URL. The URLParser removes the first
part of the URL and tries to find a matching application to relay the request
to. For instance, consider the URL \texttt{/client/file.html}. The first
part---\texttt{client}---will be removed and the request \texttt{/file.html}
will be sent to the application named \texttt{client.py}.

@@

This is a custom application built for Zohmg. It dispatches request to serve
static files from the project's \texttt{clients} directory. Custom
clients---which interact with Zohmg---can be built with HTML and JavaScript, see
the examples.

@@

The data serving part of Zohmg fetches data from HBase via its Thrift interface.
The user submits a scan query with row range, unit, and dimensions, and all the
wanted values associated with these variables.

The above query would return the pageviews for SE between 1 January
2009 and 4 April 2009. The results from the scan is served to the user as
JSONP\footnote{JSON with padding: JSON extension which adds name of callback
function as an argument to the function itself.}.

@@

Proper support for multiple data sets in a Zohmg project.

@@

\section*{Installation}

@@

\section*{Deployment}

@@

\section*{Import}

@@

\section*{Export}

@@

The above query would return the pageviews for SE between 1 January
2009 and 4 April 2009.

@@

\textbf{OLAP} &
An acronym for On-Line Analytical Processing. It is a subdivision of the broader
category \textit{business logic}. OLAP is an approach to quickly find answer to
user queries.
